Portfolio prioritization of salmon conservation

### 2013-03-12 inst/examples/salmonport.tex
Add main manuscript files in inst/examples

% Portfolio prioritization of salmon conservation
% Sean Anderson

# Abstract
First part is copied from what I wrote before (introduction and questions). Second part outlines the simulation. I haven't yet worked very much with the portfolios themselves, but I've set up the structure of the simulation. I outline how the simulation is currently working and show examples of some of the plotting functions I have to look at simulation and portfolio optimization.

# Introduction
[Based partly on Jon's notes.] Managing risk is fundamental to the conservation of an endangered species. When an endangered species exists as a metapopulation, we can manage risk at two levels: at the population level or at the metapopulation level. Typically we treat sources of risk at the metapopulation level as external and uncontrollable and so we manage risk by altering anthropogenic disturbances such as fishing or hunting on a population level as well as improving connectivity of populations.

The management of financial portfolios provides another way of considering risk. Economists consider the risk and performance of a financial portfolio as depending on the weighting of individual assets that make up the portfolio. Modern Portfolio Theory (MPT) proposes that there is a set of portfolios that achieves maximum expected return for a level of expected risk or minimum expected risk for a level of expected return [@Markowitz1952; @Markowitz1959]. This optimal set contains portfolios that range along a continuum of risk-tolerance; economists refer to this set as the efficient frontier.

Similarly, expected growth rate and variance at a metapopulation level will be a function of the properties of the populations. A portfolio approach to managing risk for a metapopulation therefore considers how conservation might affect the ``weight'' of each population in the metapopulation ``portfolio''.

@@

## Portfolio prioritization of salmon conservation
Here, we conduct a post-hoc assessment of the set of optimal salmon portfolios given various return-risk definitions and objectives. We ask: What does a portfolio approach to management tell us about salmonid conservation priorities? Does it produce any rules of thumb about what optimal portfolio management of salmonid metapopulations would entail?

## Questions
1. What are the attributes of the populations that form the optimal set? For example, are populations with high response diversity prioritized?

2. How much lower is stream-level and ESU-level extinction risk for the optimal portfolios than the less optimal portfolios? Extinction probability could be estimated through long-term simulation or through an asymmetric risk metric like conditional value at risk (CVaR). Given I am including straying, this will most likely be a quasi-extinction-risk metric.

3. How do our answers to other questions change depending on what we measure on the risk-return axes? For example, we could trade off the mean rate of change of salmon returns vs.\ variance of the rate of change. This would be most similar to a financial portfolio optimization. We could also trade off the salmon returns (or catches) themselves. Further, we could work with variance or a downside risk metric like CVaR.

4. How do constraints (e.g.\ viable salmonid population (VSP)-like constraints) affect the selected portfolios in risk-return space and in terms of common metrics such as population-level and ESU-level extinction risk?

5. How do the properties of the optimal portfolio set change based on different kinds of environmental driver patterns? For example, magnitude of driver variance, cyclical dynamics like the PDO, black swans or unexpected surprises, increases in temperature and variability through time (climate change).

Simulation
==========
See Figure \ref{fig:sim-diagram} for an illustration of the simulation 
structure.

@@

## Spawner-return relationship
I am using a Ricker curve,

@@

## Escapement targets
I am calculating spawning abundance MSY $S_{MSY}$ each generation using the equation $S_{MSY} = b(0.5-0.07a)$ from @Hilborn1992 p272, Table 7.2. I am using the precisely known $a$ and $b$ values each time and assume perfect implementation of the escapement goal (no outcome uncertainty) and no upper limit on how much the fishery can harvest. We may want to add a bit of noise here, say by assuming a constant value for $a$ or adding some implementation uncertainty. We need some level of harvesting to keep the spawning population away from carrying capacity. At carrying capacity, the productivity parameter $a$ has minimal effect on population dynamics, making it a poor way to integrate response diversity.

## Environmental signal
I have programmed five different kinds of environmental signals. I think it's easiest to think of these as temperature signals, but they could represent more general environmental anomalies, such as the PDO. The five kinds of environmental signals are described in Table \ref{tab:env-types} and illustrated in Figure \ref{fig:env-ts}. Later, I may include black-swan-like events, although the regime shifts could be seen this way if the changes are dramatic enough. So far all figures use a sine wave. The problem with using a randomly generated ARMA time series is that you get much greater differences in the portfolios from 
run to run.

@@

```{r env-ts, fig.height=5, fig.width=4, fig.cap="Example temperature time series\\label{fig:env-ts}"}
source("plot_env_ts.R")
col_pal <- rev(gg_color_hue(10))
```

## Straying
[Insert a bit here about my research on straying and it's influence on metapopulation dynamics for salmon]. Essentially, straying should be enough to prevent local extinctions and ensure all possible habitat is occupied, but probably not strong enough to noticeably affect population dynamics or synchrony [@Schtickzelle2007].

I've implemented straying as in @Cooper1999. I generate a matrix that represents the fraction of straying between any two populations. I have set up the metapopulation (thus far) as in in a very simple scenario: the populations are arranged in a line and those that are nearer to each other are more likely to stray between each other [insert caveats]. Two parameters control the straying: the fraction of fish that stray from their natal stream in any given generation $f$ and the rate at which this straying between streams decays with distance $m$. I calculated the number of salmon straying from stream $i$ to stream $j$ as,

@@

where the denominator is a normalization constant to ensure that no salmon are gained or lost. The subscript $k$ represents a stream ID. [Not sure this is written correctly --- I'll think about it --- but it should be coded correctly.] There is no straying from and to the same stream. See Figure \ref{fig:stray-matrix} for an example straying matrix.

```{r stray-matrix, fig.height=3.7, fig.width=3.9, fig.cap="Example straying matrix. Darker blue colours indicate a greater rate of straying.\\label{fig:stray-matrix}"}
source("plot_straying_matrix.R")
```

## Diversity of response to the environmental conditions
[insert research I've done here on what can generate diversity of response to the environment across populations and how it is typically incorporated] This is a critical part to the paper.

I am incorporating diversity of response to environmental conditions, say temperature, via changes in the Ricker $a$ parameter. I am using thermal tolerance curves, based loosely on @Eliason2011. They identify unique thermal tolerance curves in terms of cardiorespiratory performance. I am using approximately the same temperature ranges and variability of those curves, but applying them to productivity. I show the curves I'm currently using in Figure \ref{fig:thermal-curves}. We could vary the distribution of the optimum temperatures, the maximum $a$ values, and the width of the curves if we wanted, although I think the current curves are a simple and straightforward base scenario and we may not need to do anything more complicated.

@@

```{r thermal-curves, cache=FALSE, fig.height=3.9, fig.width=5.4, fig.cap="Thermal tolerance curves for ten populations. Here, and throughout this document, I have used the same rainbow of colours to identify the salmon populations. The populations are arranged from blue (more productive with cooler temperatures) to red (more productive with warmer temperatures\\label{fig:thermal-curves}"}
source("plot_thermal_curves.R")
```

Portfolio optimization
======================
Details to follow. Quick version: I Monte Carlo the $b$ parameters (with a specified minimum value) and measure the resulting mean and variance of the rate of growth of the metapopulation salmon returns. I then identify the set of $b$ parameters (i.e.\ the set of weights) that results in the efficient frontier. We could later choose to work with risk metrics like conditional value at risk (CVaR), which measure the probability of a ``bad'' event specifically, not the symmetrical variance.

@@

Comparing portfolios
====================
This is a big area to think about. How do portfolios differ along the frontier? How do the frontier portfolios differ from non-frontier portfolios? What is the best way to visually compare them? What aspects to compare? Are there summary statistics we could use to compare them?

Some sample simulations
=======================
The following are options and their default values:

@@

```{r run-base-example1, fig.height=6.5, fig.width=5, fig.cap="Example simulation with all parameters at default values. I'm only showing a sample of 40 generations (after the burn-in period). From top to bottom: salmon returns each generation (catch plus escapement); Ricker $a$ parameter; the environmental signal (could think of it as temperature or a composite environmental signal like the PDO); number of salmon straying by population; number of straying salmon being received by population; fisheries catch based on escapement MSY calculation; salmon escapement; residuals from the spawner-return curve."}
set.seed(2)
eg1 <- meta_sim(b = rep(1000, 10), n_pop = 10)
plot_sim_ts(eg1, years_to_show = 40)
```

```{r plot-base-corrs, fig.width=4, fig.height=4, fig.cap="A plot comparing the log(returns) between each population. The population colours and numbers match in all figures. Note how populations 1 and 10 have asynchronous returns whereas populations with more similar thermal-tolerance curves (say populations 9 and 10) have more synchronous dynamics. Populations with thermal tolerance curves in the middle (e.g.\ population 6) are less correlated with other populations. Their population dynamics end up primarily driven by demographic stochasticity and less so by temperature-induced systematic changes in productivity."}
plot_correlation_between_returns(eg1, pal = col_pal)
```

```{r plot-base-rickers, fig.width=8, fig.height=5, cache=FALSE, fig.cap="We can also look at samples of the Ricker curves by population (population numbers are in the upper left corners). Here, I've sampled 40 of the $a$ parameters through time for each population. The shaded rectangles to indicate the range of spawners observed in the simulation. The colour of the line indicates the temperature anomaly in that year with blue indicating cooler years and red indicating warmer years. In this example all the populations have the same $b$ parameters. Notice how the productivity is more variable for populations nearer the limits of their thermal tolerance (e.g.\ population 1 vs.\ population 5). Also notice how population 1 is more productive in cool years whereas population 10 is more productive in warm years."}
plot_rickers(eg1, pal = col_pal)
```

```{r plot-various-options-ts-3pops, fig.width=4, fig.height=6.3, fig.cap="Example simulations with different parameter values. I am showing three populations here to make the plots easier to interpret.\\label{fig:sim-param-ts-3}"}
source("example_returns_3.R")
```

```{r plot-various-options-ts, fig.width=4, fig.height=6.3, fig.cap="Example simulations with different parameter values. This version has ten populations.\\label{fig:sim-param-ts-10}"}
source("example_returns_10.R")
```

@@

An example of portfolio optimization
====================================
Now, let's run an example where we find the $b$ values that form the efficient frontier of metapopulation portfolios.

```{r set-a-seed, echo=FALSE}
set.seed(3)
```

```{r monte-carlo-portfolios}
weights_matrix <- create_asset_weights(n_pop = 10, n_sims = 100, 
  weight_lower_limit = 0.01)
mc_ports <- monte_carlo_portfolios(weights_matrix = weights_matrix,
  n_sims = 100, mean_b = 1000)
```

```{r eff-frontier-eg, fig.width=6.5, fig.height=3.5, fig.cap="Efficient frontier of metapopulation portfolios (red dots). Each dot represents a different set of weights of the Ricker $b$ parameters. Metapopulation growth rate is the first difference of the log of total metapopulation abundance. The colours in the right panel correspond to the other figures."}
ef_dat <- plot_efficient_portfolios(port_vals = mc_ports$port_vals, pal = 
  col_pal, weights_matrix = weights_matrix)
```

```{r plot-eff-ports, fig.height=6.5, fig.width=5, fig.cap="Simulation output from the minimum-variance metapopulation."}
min_var_sim <- mc_ports$sims_out[[ef_dat$min_var_port_id]]
plot_sim_ts(min_var_sim, pal = col_pal)
```

```{r fig.width=8, fig.height=5, fig.cap="Example Ricker curves for the minimum-variance metapopulation."}
plot_rickers(min_var_sim, pal = col_pal)
```

```{r plot-portfolio-timeseries, fig.width=6.5, fig.height=3.5, fig.cap="Time series for a random sample of portfolios (grey lines) and the minimum-variance portfolio (white lines). The top panel shows rate of change or metapopulation abundance at each time step and the bottom panel shows the metapopulation abundance."}
source("plot_portfolio_ts.R")
```

@@

References
==========

>>>>

%\documentclass[]{article}
\documentclass[12pt]{article}

% my modifications:
\usepackage{fontspec}
\usepackage{booktabs}
\defaultfontfeatures{Ligatures=TeX}
\setsansfont{Calibri}
\setmonofont{Inconsolata}
\setmainfont{Linux Libertine}
\DeclareTextCommandDefault{\nobreakspace}{\leavevmode\nobreak\ }
\widowpenalty=1000
\clubpenalty=1000
\usepackage{setspace}
\onehalfspacing

% A bit bigger than one inch margins:
%\topmargin -1.43cm
%\oddsidemargin 0.4cm
%\evensidemargin 0.4cm
%\textwidth 14.5cm
\textheight 22.5cm

%\usepackage[left=1.5in,top=1.5in,right=1.5in,bottom=1.5in,nohead,nofoot]{geometry}
\usepackage{geometry}
\geometry{letterpaper}

\usepackage[round]{natbib}
\bibliographystyle{cjfas}
\bibpunct{(}{)}{;}{a}{}{;}

%\usepackage[T1]{fontenc}
%\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
%\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
%\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  %\usepackage[utf8]{inputenc}
%%\else % if luatex or xelatex
  %\usepackage{fontspec}
  %\ifxetex
    %\usepackage{xltxtra,xunicode}
  %\fi
  %\defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  %\newcommand{\euro}{€}
%%%%%\fi
% use microtype if available
%\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
%%%\usepackage{natbib}
%\bibliographystyle{plainnat}
%%%%%\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
%\makeatletter
%\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
%\else\Gin@nat@width\fi}
%\makeatother
%\let\Oldincludegraphics\includegraphics
%\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
%\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
%
\title{Portfolio prioritization of salmon conservation}
\author{Sean Anderson}
\date{}

\begin{document}
\maketitle

\section{Abstract}

First part is copied from what I wrote before (introduction and
questions). Second part outlines the simulation. I haven't yet worked
very much with the portfolios themselves, but I've set up the structure
of the simulation. I outline how the simulation is currently working and
show examples of some of the plotting functions I have to look at
simulation and portfolio optimization.

\section{Introduction}

{[}Based partly on Jon's notes.{]} Managing risk is fundamental to the
conservation of an endangered species. When an endangered species exists
as a metapopulation, we can manage risk at two levels: at the population
level or at the metapopulation level. Typically we treat sources of risk
at the metapopulation level as external and uncontrollable and so we
manage risk by altering anthropogenic disturbances such as fishing or
hunting on a population level as well as improving connectivity of
populations.

The management of financial portfolios provides another way of
considering risk. Economists consider the risk and performance of a
financial portfolio as depending on the weighting of individual assets
that make up the portfolio. Modern Portfolio Theory (MPT) proposes that
there is a set of portfolios that achieves maximum expected return for a
level of expected risk or minimum expected risk for a level of expected
return \citep{Markowitz1952, Markowitz1959}. This optimal set contains
portfolios that range along a continuum of risk-tolerance; economists
refer to this set as the efficient frontier.

Similarly, expected growth rate and variance at a metapopulation level
will be a function of the properties of the populations. A portfolio
approach to managing risk for a metapopulation therefore considers how
conservation might affect the
\texttt{weight'' of each population in the metapopulation}portfolio''.

@@

\subsection{Portfolio prioritization of salmon conservation}

Here, we conduct a post-hoc assessment of the set of optimal salmon
portfolios given various return-risk definitions and objectives. We ask:
What does a portfolio approach to management tell us about salmonid
conservation priorities? Does it produce any rules of thumb about what
optimal portfolio management of salmonid metapopulations would entail?

\subsection{Questions}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  What are the attributes of the populations that form the optimal set?
  For example, are populations with high response diversity prioritized?
\item
  How much lower is stream-level and ESU-level extinction risk for the
  optimal portfolios than the less optimal portfolios? Extinction
  probability could be estimated through long-term simulation or through
  an asymmetric risk metric like conditional value at risk (CVaR). Given
  I am including straying, this will most likely be a
  quasi-extinction-risk metric.
\item
  How do our answers to other questions change depending on what we
  measure on the risk-return axes? For example, we could trade off the
  mean rate of change of salmon returns vs.~variance of the rate of
  change. This would be most similar to a financial portfolio
  optimization. We could also trade off the salmon returns (or catches)
  themselves. Further, we could work with variance or a downside risk
  metric like CVaR.
\item
  How do constraints (e.g.~viable salmonid population (VSP)-like
  constraints) affect the selected portfolios in risk-return space and
  in terms of common metrics such as population-level and ESU-level
  extinction risk?
\item
  How do the properties of the optimal portfolio set change based on
  different kinds of environmental driver patterns? For example,
  magnitude of driver variance, cyclical dynamics like the PDO, black
  swans or unexpected surprises, increases in temperature and
  variability through time (climate change).
\end{enumerate}

\section{Simulation}

See Figure \ref{fig:sim-diagram} for an illustration of the simulation
structure.

@@

\subsection{Spawner-return relationship}

I am using a Ricker curve,

@@

\subsection{Escapement targets}

I am calculating spawning abundance MSY $S_{MSY}$ each generation using
the equation $S_{MSY} = b(0.5-0.07a)$ from \citet{Hilborn1992} p272,
Table 7.2. I am using the precisely known $a$ and $b$ values each time
and assume perfect implementation of the escapement goal (no outcome
uncertainty) and no upper limit on how much the fishery can harvest. We
may want to add a bit of noise here, say by assuming a constant value
for $a$ or adding some implementation uncertainty. We need some level of
harvesting to keep the spawning population away from carrying capacity.
At carrying capacity, the productivity parameter $a$ has minimal effect
on population dynamics, making it a poor way to integrate response
diversity.

\subsection{Environmental signal}

I have programmed five different kinds of environmental signals. I think
it's easiest to think of these as temperature signals, but they could
represent more general environmental anomalies, such as the PDO. The
five kinds of environmental signals are described in Table
\ref{tab:env-types} and illustrated in Figure \ref{fig:env-ts}. Later, I
may include black-swan-like events, although the regime shifts could be
seen this way if the changes are dramatic enough. So far all figures use
a sine wave. The problem with using a randomly generated ARMA time
series is that you get much greater differences in the portfolios from
run to run.

@@

\begin{figure}[htbp]
\centering
\includegraphics{figure/env-ts.pdf}
\caption{Example temperature time series\label{fig:env-ts}}
\end{figure}

\subsection{Straying}

{[}Insert a bit here about my research on straying and it's influence on
metapopulation dynamics for salmon{]}. Essentially, straying should be
enough to prevent local extinctions and ensure all possible habitat is
occupied, but probably not strong enough to noticeably affect population
dynamics or synchrony \citep{Schtickzelle2007}.

I've implemented straying as in \citet{Cooper1999}. I generate a matrix
that represents the fraction of straying between any two populations. I
have set up the metapopulation (thus far) as in in a very simple
scenario: the populations are arranged in a line and those that are
nearer to each other are more likely to stray between each other
{[}insert caveats{]}. Two parameters control the straying: the fraction
of fish that stray from their natal stream in any given generation $f$
and the rate at which this straying between streams decays with distance
$m$. I calculated the number of salmon straying from stream $i$ to
stream $j$ as,

@@

where the denominator is a normalization constant to ensure that no
salmon are gained or lost. The subscript $k$ represents a stream ID.
{[}Not sure this is written correctly --- I'll think about it --- but it
should be coded correctly.{]} There is no straying from and to the same
stream. See Figure \ref{fig:stray-matrix} for an example straying
matrix.

\begin{figure}[htbp]
\centering
\includegraphics{figure/stray-matrix.pdf}
\caption{Example straying matrix. Darker blue colours indicate a greater
rate of straying.\label{fig:stray-matrix}}
\end{figure}

\subsection{Diversity of response to the environmental conditions}

{[}insert research I've done here on what can generate diversity of
response to the environment across populations and how it is typically
incorporated{]} This is a critical part to the paper.

I am incorporating diversity of response to environmental conditions,
say temperature, via changes in the Ricker $a$ parameter. I am using
thermal tolerance curves, based loosely on \citet{Eliason2011}. They
identify unique thermal tolerance curves in terms of cardiorespiratory
performance. I am using approximately the same temperature ranges and
variability of those curves, but applying them to productivity. I show
the curves I'm currently using in Figure \ref{fig:thermal-curves}. We
could vary the distribution of the optimum temperatures, the maximum $a$
values, and the width of the curves if we wanted, although I think the
current curves are a simple and straightforward base scenario and we may
not need to do anything more complicated.

@@

\begin{figure}[htbp]
\centering
\includegraphics{figure/thermal-curves.pdf}
\caption{Thermal tolerance curves for ten populations. Here, and
throughout this document, I have used the same rainbow of colours to
identify the salmon populations. The populations are arranged from blue
(more productive with cooler temperatures) to red (more productive with
warmer temperatures\label{fig:thermal-curves}}
\end{figure}

\section{Portfolio optimization}

Details to follow. Quick version: I Monte Carlo the $b$ parameters (with
a specified minimum value) and measure the resulting mean and variance
of the rate of growth of the metapopulation salmon returns. I then
identify the set of $b$ parameters (i.e.~the set of weights) that
results in the efficient frontier. We could later choose to work with
risk metrics like conditional value at risk (CVaR), which measure the
probability of a ``bad'' event specifically, not the symmetrical
variance.

@@

\section{Comparing portfolios}

This is a big area to think about. How do portfolios differ along the
frontier? How do the frontier portfolios differ from non-frontier
portfolios? What is the best way to visually compare them? What aspects
to compare? Are there summary statistics we could use to compare them?

\section{Some sample simulations}

The following are options and their default values:

@@

\begin{figure}[htbp]
\centering
\includegraphics{figure/run-base-example1.pdf}
\caption{Example simulation with all parameters at default values. I'm
only showing a sample of 40 generations (after the burn-in period). From
top to bottom: salmon returns each generation (catch plus escapement);
Ricker $a$ parameter; the environmental signal (could think of it as
temperature or a composite environmental signal like the PDO); number of
salmon straying by population; number of straying salmon being received
by population; fisheries catch based on escapement MSY calculation;
salmon escapement; residuals from the spawner-return curve.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figure/plot-base-corrs.pdf}
\caption{A plot comparing the log(returns) between each population. The
population colours and numbers match in all figures. Note how
populations 1 and 10 have asynchronous returns whereas populations with
more similar thermal-tolerance curves (say populations 9 and 10) have
more synchronous dynamics. Populations with thermal tolerance curves in
the middle (e.g.~population 6) are less correlated with other
populations. Their population dynamics end up primarily driven by
demographic stochasticity and less so by temperature-induced systematic
changes in productivity.}
\end{figure}

\begin{verbatim}
Error: could not find function "ricker_ar1"
\end{verbatim}

\begin{figure}[htbp]
\centering
\includegraphics{figure/plot-various-options-ts-3pops.pdf}
\caption{Example simulations with different parameter values. I am
showing three populations here to make the plots easier to
interpret.\label{fig:sim-param-ts-3}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figure/plot-various-options-ts.pdf}
\caption{Example simulations with different parameter values. This
version has ten populations.\label{fig:sim-param-ts-10}}
\end{figure}

@@

\section{An example of portfolio optimization}

Now, let's run an example where we find the $b$ values that form the
efficient frontier of metapopulation portfolios.

\begin{figure}[htbp]
\centering
\includegraphics{figure/eff-frontier-eg.pdf}
\caption{Efficient frontier of metapopulation portfolios (red dots).
Each dot represents a different set of weights of the Ricker $b$
parameters. Metapopulation growth rate is the first difference of the
log of total metapopulation abundance. The colours in the right panel
correspond to the other figures.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figure/plot-eff-ports.pdf}
\caption{Simulation output from the minimum-variance metapopulation.}
\end{figure}

\begin{verbatim}
Error: could not find function "ricker_ar1"
\end{verbatim}

\begin{figure}[htbp]
\centering
\includegraphics{figure/plot-portfolio-timeseries.pdf}
\caption{Time series for a random sample of portfolios (grey lines) and
the minimum-variance portfolio (white lines). The top panel shows rate
of change or metapopulation abundance at each time step and the bottom
panel shows the metapopulation abundance.}
\end{figure}

@@

\renewcommand\refname{References}
\bibliography{jshort,salmonport}

\end{document}
